---
title: "Vignette Project"
author: "Jessica Ayers"
date: "2023-06-22"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(fig.path = "/Users/jessayers/Documents/ST 558/TOPIC 2/Project-1/")
```

# Packages Needed

The packages needed for analysis of the Open Movie Database API are:

  - `httr`
  - `tidyverse`
  - `jsonlite`
  - `GGally`

```{r, message = FALSE}
library(httr)
library(jsonlite)
library(tidyverse)
library(GGally)
```

# Functions

The API used was the Open Movie Database API. Two different types of data were returned based on parameters used in the link. With the below function, a list of information corresponding to one random movie will returned along with a data frame of 10 random movies with more limited information. The data frame will be used to do the exploratory data analysis. For this API, movies with different colors in the titles will be returned for three different release years. The 6 choices are:

  - Options for `s` include *blue*, *red* and *green*
    - `s` searches for movie titles with the provided word
  - Options for `y` include *2000*, *2010* and *2020*
    - `y` stands for year of release
    
There are 9 total combinations of possible modifications. 

```{r}
contactFunction <- function(s, y){

if((s == "blue") & (y == 2000)) {

blueData <- GET("http://www.omdbapi.com/?s=blue&y=2000&apikey=ca153ce0")
blueParsed <- fromJSON(rawToChar(blueData$content))

bluedf <- data.frame(blueParsed$Search$Title, blueParsed$Search$Year, blueParsed$Search$imdbID, blueParsed$Search$Type)
blue <- bluedf %>%
  as_tibble() %>%
  rename(Title = blueParsed.Search.Title, Year = blueParsed.Search.Year, imdbID = blueParsed.Search.imdbID, Type = blueParsed.Search.Type)

testData <- GET("http://www.omdbapi.com/?t=blue&y=2000&apikey=ca153ce0")
randomData <- fromJSON(rawToChar(testData$content))

return(list(RandomMovie = randomData, RandomDF = blue))
}

else if((s == "blue") & (y == 2010)){
  blueData <- GET("http://www.omdbapi.com/?s=blue&y=2010&apikey=ca153ce0")
  blueParsed <- fromJSON(rawToChar(blueData$content))

  bluedf <- data.frame(blueParsed$Search$Title, blueParsed$Search$Year, blueParsed$Search$imdbID, blueParsed$Search$Type)
  blue <- bluedf %>%
    as_tibble() %>%
    rename(Title = blueParsed.Search.Title, Year = blueParsed.Search.Year, imdbID = blueParsed.Search.imdbID, Type = blueParsed.Search.Type) 

testData <- GET("http://www.omdbapi.com/?t=blue&y=2010&apikey=ca153ce0")
randomData <- fromJSON(rawToChar(testData$content))

return(list(RandomMovie = randomData, RandomDF = blue))
}

else if((s == "blue") & (y == 2020)){
  blueData <- GET("http://www.omdbapi.com/?s=blue&y=2020&apikey=ca153ce0")
blueParsed <- fromJSON(rawToChar(blueData$content))

bluedf <- data.frame(blueParsed$Search$Title, blueParsed$Search$Year, blueParsed$Search$imdbID, blueParsed$Search$Type)
blue <- bluedf %>%
  as_tibble() %>%
  rename(Title = blueParsed.Search.Title, Year = blueParsed.Search.Year, imdbID = blueParsed.Search.imdbID, Type = blueParsed.Search.Type)

testData <- GET("http://www.omdbapi.com/?t=blue&y=2020&apikey=ca153ce0")
randomData <- fromJSON(rawToChar(testData$content))

return(list(RandomMovie = randomData, RandomDF = blue))
}

else if ((s == "red") & (y == 2000)){
  redData <- GET("http://www.omdbapi.com/?s=red&y=2000&apikey=ca153ce0")
redParsed <- fromJSON(rawToChar(redData$content))

reddf <- data.frame(redParsed$Search$Title, redParsed$Search$Year, redParsed$Search$imdbID, redParsed$Search$Type)
red <- reddf %>%
  as_tibble() %>%
  rename(Title = redParsed.Search.Title, Year = redParsed.Search.Year, imdbID = redParsed.Search.imdbID, Type = redParsed.Search.Type)

testData <- GET("http://www.omdbapi.com/?t=red&y=2000&apikey=ca153ce0")
randomData <- fromJSON(rawToChar(testData$content))

return(list(RandomMovie = randomData, RandomDF = red))
}

else if ((s == "red") & (y == 2010)){
  redData <- GET("http://www.omdbapi.com/?s=red&y=2010&apikey=ca153ce0")
redParsed <- fromJSON(rawToChar(redData$content))

reddf <- data.frame(redParsed$Search$Title, redParsed$Search$Year, redParsed$Search$imdbID, redParsed$Search$Type)
red <- reddf %>%
  as_tibble() %>%
  rename(Title = redParsed.Search.Title, Year = redParsed.Search.Year, imdbID = redParsed.Search.imdbID, Type = redParsed.Search.Type) 

testData <- GET("http://www.omdbapi.com/?t=red&y=2010&apikey=ca153ce0")
randomData <- fromJSON(rawToChar(testData$content))

return(list(RandomMovie = randomData, RandomDF = red))
}

else if((s == "red") & (y == 2020)){
  redData <- GET("http://www.omdbapi.com/?s=red&y=2020&apikey=ca153ce0")
redParsed <- fromJSON(rawToChar(redData$content))

reddf <- data.frame(redParsed$Search$Title, redParsed$Search$Year, redParsed$Search$imdbID, redParsed$Search$Type)
red <- reddf %>%
  as_tibble() %>%
  rename(Title = redParsed.Search.Title, Year = redParsed.Search.Year, imdbID = redParsed.Search.imdbID, Type = redParsed.Search.Type) 

testData <- GET("http://www.omdbapi.com/?t=red&y=2020&apikey=ca153ce0")
randomData <- fromJSON(rawToChar(testData$content))

return(list(RandomMovie = randomData, RandomDF = red))
}

else if((s == "green") & (y == 2000)){
  greenData <- GET("http://www.omdbapi.com/?s=green&y=2000&apikey=ca153ce0")
greenParsed <- fromJSON(rawToChar(greenData$content))

greendf <- data.frame(greenParsed$Search$Title, greenParsed$Search$Year, greenParsed$Search$imdbID, greenParsed$Search$Type)
green <- greendf %>%
  as_tibble() %>%
  rename(Title = greenParsed.Search.Title, Year = greenParsed.Search.Year, imdbID = greenParsed.Search.imdbID, Type = greenParsed.Search.Type) 

testData <- GET("http://www.omdbapi.com/?t=green&y=2000&apikey=ca153ce0")
randomData <- fromJSON(rawToChar(testData$content))

return(list(RandomMovie = randomData, RandomDF = green))
}

else if((s== "green") & (y == 2010)){
  greenData <- GET("http://www.omdbapi.com/?s=green&y=2010&apikey=ca153ce0")
greenParsed <- fromJSON(rawToChar(greenData$content))

greendf <- data.frame(greenParsed$Search$Title, greenParsed$Search$Year, greenParsed$Search$imdbID, greenParsed$Search$Type)
green <- greendf %>%
  as_tibble() %>%
  rename(Title = greenParsed.Search.Title, Year = greenParsed.Search.Year, imdbID = greenParsed.Search.imdbID, Type = greenParsed.Search.Type) 

testData <- GET("http://www.omdbapi.com/?t=green&y=2010&apikey=ca153ce0")
randomData <- fromJSON(rawToChar(testData$content))
return(list(RandomMovie = randomData, RandomDF = green))
}

else if((s == "green") & (y == 2020)){
  greenData <- GET("http://www.omdbapi.com/?s=green&y=2020&apikey=ca153ce0")
greenParsed <- fromJSON(rawToChar(greenData$content))

greendf <- data.frame(greenParsed$Search$Title, greenParsed$Search$Year, greenParsed$Search$imdbID, greenParsed$Search$Type)
green <- greendf %>%
  as_tibble() %>%
  rename(Title = greenParsed.Search.Title, Year = greenParsed.Search.Year, imdbID = greenParsed.Search.imdbID, Type = greenParsed.Search.Type) 

testData <- GET("http://www.omdbapi.com/?t=green&y=2020&apikey=ca153ce0")
randomData <- fromJSON(rawToChar(testData$content))

return(list(RandomMovie = randomData, RandomDF = green))
}

else(stop("invalid s or y argument"))

}
```

Some examples of random movie titles returned from the above function are:
```{r}
contactFunction("blue", 2010)$RandomMovie
contactFunction("red", 2010)$RandomMovie
contactFunction("green", 2010)$RandomMovie
```

Examples of data frames that are returned from the above function:

```{r}
contactFunction("blue", 2010)$RandomDF
contactFunction("red", 2010)$RandomDF
contactFunction("green", 2010)$RandomDF
```

# Exploratory Data Analysis

```{r, message = FALSE}
redData1 <- contactFunction(s = "red", y = 2000)$RandomDF
redData2 <- contactFunction(s = "red", y = 2010)$RandomDF
redData3 <- contactFunction(s = "red", y = 2020)$RandomDF
half <- full_join(redData1, redData2)
edaData <- full_join(half, redData3)
edaData
```

```{r}
#Create new variable
#initialize Sequel Variable
edaData$Sequel <- ""
for(i in 1:30){
  if(grepl("2", edaData$Title[i]) == "TRUE"){
    edaData$Sequel[i] = "Yes"
  }
  else {
    edaData$Sequel[i] = "No"
  }
}
edaData

#initialize length of title variable
edaData$lengthOfTitle <- 0
for(i in 1:30){
  edaData$lengthOfTitle[i] <- nchar(edaData$Title[i])
}
edaData
```

```{r}
#contingency tables
table(edaData$Year)
table(edaData$Type)
table(edaData$Type, edaData$Year)
table(edaData$Type, edaData$Year, edaData$Sequel)
```

```{r}
#numerical summaries for quantitative variables at each setting of the categorical variables
mean(edaData$lengthOfTitle)
median(edaData$lengthOfTitle)
sd(edaData$lengthOfTitle)
var(edaData$lengthOfTitle)
IQR(edaData$lengthOfTitle)

#each setting of year first
edaData %>% group_by(Year) %>%
summarise(avg = mean(lengthOfTitle), med = median(lengthOfTitle), var = var(lengthOfTitle))

#each setting of type
edaData %>% group_by(Type) %>%
summarise(avg = mean(lengthOfTitle), med = median(lengthOfTitle), var = var(lengthOfTitle))

#each setting of Sequel
edaData %>% group_by(Sequel) %>%
summarise(avg = mean(lengthOfTitle), med = median(lengthOfTitle), var = var(lengthOfTitle))
```

```{r}
#bar plot
g <- ggplot(data = edaData, aes(x = Year))
g + geom_bar(aes(fill = Type)) + 
  labs(title = "Bar Plot of Years of Movies, Series, and Games") +
  coord_flip()
```

```{r}
#histogram
g <- ggplot(edaData, aes(x = lengthOfTitle))
g + geom_histogram(color = "black", fill = "white",
size = 2, binwidth = 3) + 
  labs(title = "Histogram of Length of Title", x = "Length of Title", y = "Count")
```

```{r}
#Boxplot
g <- ggplot(edaData, aes(x = Type, y = lengthOfTitle))
g + geom_boxplot(fill = "red")
```
```{r}
g <- ggplot(edaData, aes(x = Year, y = lengthOfTitle))
g + geom_boxplot(fill = "red") + 
stat_summary(fun.y = median, geom = "line",
lwd = 1.5, aes(group = Type, col = Type))
```

```{r}
#scatterplot
g <- ggplot(edaData, aes(x = lengthOfTitle, y = imdbID))
g + geom_point() +
  facet_grid(~edaData$Type) + 
  geom_text(aes(label = Year))
```

```{r}
g <- ggplot(edaData, aes(x = Sequel, y = lengthOfTitle, color = Year))
g + geom_line(lwd = 4)
```


