---
title: "Vignette Project"
author: "Jessica Ayers"
date: "2023-06-22"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(fig.path = "/Users/jessayers/Documents/ST 558/TOPIC 2/Project-1/")
```

# Packages Needed

The packages needed for analysis of the Open Movie Database API are:

  - `httr`
  - `tidyverse`
  - `jsonlite`
  - `GGally`

```{r, message = FALSE}
library(httr)
library(jsonlite)
library(tidyverse)
library(GGally)
```

# Functions

The API used was the Open Movie Database API. Two different types of data were returned based on parameters used in the link. With the below function, a list of information corresponding to one random movie will returned along with a data frame of 10 random movies with more limited information. The data frame will be used to do the exploratory data analysis. For this API, movies with different colors in the titles will be returned for three different release years. The 6 choices are:

  - Options for `s` include *blue*, *red* and *green*
    - `s` searches for movie titles with the provided word
  - Options for `y` include *2000*, *2010* and *2020*
    - `y` stands for year of release
    
There are 9 total combinations of possible modifications. 

```{r}
contactFunction <- function(s, y){

if((s == "blue") & (y == 2000)) {

blueData <- GET("http://www.omdbapi.com/?s=blue&y=2000&apikey=ca153ce0")
blueParsed <- fromJSON(rawToChar(blueData$content))

bluedf <- data.frame(blueParsed$Search$Title, blueParsed$Search$Year, blueParsed$Search$imdbID, blueParsed$Search$Type)
blue <- bluedf %>%
  as_tibble() %>%
  rename(Title = blueParsed.Search.Title, Year = blueParsed.Search.Year, imdbID = blueParsed.Search.imdbID, Type = blueParsed.Search.Type)

testData <- GET("http://www.omdbapi.com/?t=blue&y=2000&apikey=ca153ce0")
randomData <- fromJSON(rawToChar(testData$content))

return(list(RandomMovie = randomData, RandomDF = blue))
}

else if((s == "blue") & (y == 2010)){
  blueData <- GET("http://www.omdbapi.com/?s=blue&y=2010&apikey=ca153ce0")
  blueParsed <- fromJSON(rawToChar(blueData$content))

  bluedf <- data.frame(blueParsed$Search$Title, blueParsed$Search$Year, blueParsed$Search$imdbID, blueParsed$Search$Type)
  blue <- bluedf %>%
    as_tibble() %>%
    rename(Title = blueParsed.Search.Title, Year = blueParsed.Search.Year, imdbID = blueParsed.Search.imdbID, Type = blueParsed.Search.Type) 

testData <- GET("http://www.omdbapi.com/?t=blue&y=2010&apikey=ca153ce0")
randomData <- fromJSON(rawToChar(testData$content))

return(list(RandomMovie = randomData, RandomDF = blue))
}

else if((s == "blue") & (y == 2020)){
  blueData <- GET("http://www.omdbapi.com/?s=blue&y=2020&apikey=ca153ce0")
blueParsed <- fromJSON(rawToChar(blueData$content))

bluedf <- data.frame(blueParsed$Search$Title, blueParsed$Search$Year, blueParsed$Search$imdbID, blueParsed$Search$Type)
blue <- bluedf %>%
  as_tibble() %>%
  rename(Title = blueParsed.Search.Title, Year = blueParsed.Search.Year, imdbID = blueParsed.Search.imdbID, Type = blueParsed.Search.Type)

testData <- GET("http://www.omdbapi.com/?t=blue&y=2020&apikey=ca153ce0")
randomData <- fromJSON(rawToChar(testData$content))

return(list(RandomMovie = randomData, RandomDF = blue))
}

else if ((s == "red") & (y == 2000)){
  redData <- GET("http://www.omdbapi.com/?s=red&y=2000&apikey=ca153ce0")
redParsed <- fromJSON(rawToChar(redData$content))

reddf <- data.frame(redParsed$Search$Title, redParsed$Search$Year, redParsed$Search$imdbID, redParsed$Search$Type)
red <- reddf %>%
  as_tibble() %>%
  rename(Title = redParsed.Search.Title, Year = redParsed.Search.Year, imdbID = redParsed.Search.imdbID, Type = redParsed.Search.Type)

testData <- GET("http://www.omdbapi.com/?t=red&y=2000&apikey=ca153ce0")
randomData <- fromJSON(rawToChar(testData$content))

return(list(RandomMovie = randomData, RandomDF = red))
}

else if ((s == "red") & (y == 2010)){
  redData <- GET("http://www.omdbapi.com/?s=red&y=2010&apikey=ca153ce0")
redParsed <- fromJSON(rawToChar(redData$content))

reddf <- data.frame(redParsed$Search$Title, redParsed$Search$Year, redParsed$Search$imdbID, redParsed$Search$Type)
red <- reddf %>%
  as_tibble() %>%
  rename(Title = redParsed.Search.Title, Year = redParsed.Search.Year, imdbID = redParsed.Search.imdbID, Type = redParsed.Search.Type) 

testData <- GET("http://www.omdbapi.com/?t=red&y=2010&apikey=ca153ce0")
randomData <- fromJSON(rawToChar(testData$content))

return(list(RandomMovie = randomData, RandomDF = red))
}

else if((s == "red") & (y == 2020)){
  redData <- GET("http://www.omdbapi.com/?s=red&y=2020&apikey=ca153ce0")
redParsed <- fromJSON(rawToChar(redData$content))

reddf <- data.frame(redParsed$Search$Title, redParsed$Search$Year, redParsed$Search$imdbID, redParsed$Search$Type)
red <- reddf %>%
  as_tibble() %>%
  rename(Title = redParsed.Search.Title, Year = redParsed.Search.Year, imdbID = redParsed.Search.imdbID, Type = redParsed.Search.Type) 

testData <- GET("http://www.omdbapi.com/?t=red&y=2020&apikey=ca153ce0")
randomData <- fromJSON(rawToChar(testData$content))

return(list(RandomMovie = randomData, RandomDF = red))
}

else if((s == "green") & (y == 2000)){
  greenData <- GET("http://www.omdbapi.com/?s=green&y=2000&apikey=ca153ce0")
greenParsed <- fromJSON(rawToChar(greenData$content))

greendf <- data.frame(greenParsed$Search$Title, greenParsed$Search$Year, greenParsed$Search$imdbID, greenParsed$Search$Type)
green <- greendf %>%
  as_tibble() %>%
  rename(Title = greenParsed.Search.Title, Year = greenParsed.Search.Year, imdbID = greenParsed.Search.imdbID, Type = greenParsed.Search.Type) 

testData <- GET("http://www.omdbapi.com/?t=green&y=2000&apikey=ca153ce0")
randomData <- fromJSON(rawToChar(testData$content))

return(list(RandomMovie = randomData, RandomDF = green))
}

else if((s== "green") & (y == 2010)){
  greenData <- GET("http://www.omdbapi.com/?s=green&y=2010&apikey=ca153ce0")
greenParsed <- fromJSON(rawToChar(greenData$content))

greendf <- data.frame(greenParsed$Search$Title, greenParsed$Search$Year, greenParsed$Search$imdbID, greenParsed$Search$Type)
green <- greendf %>%
  as_tibble() %>%
  rename(Title = greenParsed.Search.Title, Year = greenParsed.Search.Year, imdbID = greenParsed.Search.imdbID, Type = greenParsed.Search.Type) 

testData <- GET("http://www.omdbapi.com/?t=green&y=2010&apikey=ca153ce0")
randomData <- fromJSON(rawToChar(testData$content))
return(list(RandomMovie = randomData, RandomDF = green))
}

else if((s == "green") & (y == 2020)){
  greenData <- GET("http://www.omdbapi.com/?s=green&y=2020&apikey=ca153ce0")
greenParsed <- fromJSON(rawToChar(greenData$content))

greendf <- data.frame(greenParsed$Search$Title, greenParsed$Search$Year, greenParsed$Search$imdbID, greenParsed$Search$Type)
green <- greendf %>%
  as_tibble() %>%
  rename(Title = greenParsed.Search.Title, Year = greenParsed.Search.Year, imdbID = greenParsed.Search.imdbID, Type = greenParsed.Search.Type) 

testData <- GET("http://www.omdbapi.com/?t=green&y=2020&apikey=ca153ce0")
randomData <- fromJSON(rawToChar(testData$content))

return(list(RandomMovie = randomData, RandomDF = green))
}

else(stop("invalid s or y argument"))

}
```

Some examples of random movie titles returned from the above function are:
```{r}
contactFunction("blue", 2010)$RandomMovie
contactFunction("red", 2010)$RandomMovie
contactFunction("green", 2010)$RandomMovie
```

Examples of data frames that are returned from the above function:

```{r}
contactFunction("blue", 2010)$RandomDF
contactFunction("red", 2010)$RandomDF
contactFunction("green", 2010)$RandomDF
```

# Exploratory Data Analysis

An exploratory data analysis was done using three endpoints. Three data frames are returned for movie titles containing the word "red" for years 2000, 2010, and 2020. A combined data frame was created and printed.

```{r, message = FALSE}
redData1 <- contactFunction(s = "red", y = 2000)$RandomDF
redData2 <- contactFunction(s = "red", y = 2010)$RandomDF
redData3 <- contactFunction(s = "red", y = 2020)$RandomDF
half <- full_join(redData1, redData2)
edaData <- full_join(half, redData3)
edaData
```

A new character variable was created to determine if a movie was a sequel or not. In addition, since all character data was returned from the API, a numeric variable was created from the length of each program title. An updated tibble was printed.

```{r}
#Create new variable
#initialize Sequel Variable
edaData$Sequel <- ""
for(i in 1:30){
  #greply used to determine if string contains a 2
  if(grepl("2", edaData$Title[i]) == "TRUE"){
    edaData$Sequel[i] = "Yes"
  }
  else {
    edaData$Sequel[i] = "No"
  }
}

#initialize length of title variable
edaData$lengthOfTitle <- 0
for(i in 1:30){
  #nchar used to determine length of string
  edaData$lengthOfTitle[i] <- nchar(edaData$Title[i])
}
edaData
```

A one-way contingency table was created for each qualitative variable.  

```{r}
#contingency tables
table(edaData$Year)
table(edaData$Type)
table(edaData$Sequel)
```

From this data, two movies were sequels. There were 3 games, 24 movies, and 3 series. 

A two-way contingency table was created for Type and Year.

```{r}
table(edaData$Type, edaData$Year)
```
There was 1 game released in 2000, 2 in 2010, and 0 in 2020. 

A three-way table was created for Type, Year, and Sequel.

```{r}
table(edaData$Type, edaData$Year, edaData$Sequel)
```
For programs that are sequels, there was 1 game and 1 movie, both released in 2000. 

Next, numerical summaries were created for the lengthOfTitle quantitative variable. 

```{r}
#numerical summaries for quantitative variables at each setting of the categorical variables
mean(edaData$lengthOfTitle)
median(edaData$lengthOfTitle)
sd(edaData$lengthOfTitle)
var(edaData$lengthOfTitle)
IQR(edaData$lengthOfTitle)
```
The mean length of title from this data was `r mean(edaData$lengthOfTitle)`. The median length of title was `r median(edaData$lengthOfTitle)`. More summary statistics are given above. 

The mean, median, and variance of length of title were then found for each grouping of Year, Type, and Sequel.

```{r}
#each setting of year first
edaData %>% group_by(Year) %>%
summarise(avg = mean(lengthOfTitle), med = median(lengthOfTitle), var = var(lengthOfTitle))

#each setting of type
edaData %>% group_by(Type) %>%
summarise(avg = mean(lengthOfTitle), med = median(lengthOfTitle), var = var(lengthOfTitle))

#each setting of Sequel
edaData %>% group_by(Sequel) %>%
summarise(avg = mean(lengthOfTitle), med = median(lengthOfTitle), var = var(lengthOfTitle))
```
NA's were created as a result of only one program being released during a series of years. 

A bar plot was created to show the count of programs per year/year interval at each type level.

```{r}
#bar plot
g <- ggplot(data = edaData, aes(x = Year))
g + geom_bar(aes(fill = Type)) + 
  labs(title = "Bar Plot of Years of Movies, Series, and Games") +
  coord_flip()
```

From the above plot, no games were released during 2020. The most games were released in 2010. In addition, no tv series were released in 2010. All three years had the same average number of releases for movies. 

Next a histogram was created.

```{r}
#histogram
g <- ggplot(edaData, aes(x = lengthOfTitle))
g + geom_histogram(color = "black", fill = "white",
size = 2, binwidth = 3) + 
  labs(title = "Histogram of Length of Title", x = "Length of Title", y = "Count")
```

The above histogram shows that less programs had longer title lengths. Most title lengths were between 5 and 20 letters long. 

A boxplot for each type was created. 

```{r}
#Boxplot
g <- ggplot(edaData, aes(x = Type, y = lengthOfTitle))
g + geom_boxplot(fill = "red") + 
  labs(y = "Length of Title", title = "Boxplot for Length of Title at Each Type Level")
```

The average length of title for games was larger than movies and series. The average length of title appears to be equivalent for movies and series, with movies have a noticeable outlying value. 

More boxplots were created for each year with the median trend line shown for each type. 

```{r}
g <- ggplot(edaData, aes(x = Year, y = lengthOfTitle))
g + geom_boxplot(fill = "red") + 
stat_summary(fun.y = median, geom = "line",
lwd = 1.5, aes(group = Type, col = Type)) + 
  labs(y = "Length of Tile", title = "Boxplots for Year with Median Lines")
```

The median length of title increases for movies from 2000 to 2020. The median length of title decreases slightly for games from 2000 to 2010. The median length of title for series fluctuates between the interval of years.

A scatterplot was created for length of title per each imdbID.

```{r}
#scatterplot
g <- ggplot(edaData, aes(x = lengthOfTitle, y = imdbID))
g + geom_point() +
  facet_grid(~edaData$Type) + 
  geom_text(aes(label = Year)) + 
  labs(y = "imdbID", x = "Length of Title", title = "Scatterplot of Length of Title for Each IMDBID")
```

The majority of imdbID's corresponded to movie titles. Most movie titles had title lengths with less than 20 characters. Each imdbID is labeled with the release year. 

Lastly, a line plot was created to show length of title for sequels vs non-sequels.

```{r}
g <- ggplot(edaData, aes(x = Sequel, y = lengthOfTitle, color = Year))
g + geom_line(lwd = 4) + 
  labs(y = "Length of Title", title = "Length of Title for Sequels vs Non-Sequels")
```

From the above plot, sequels were found in only 2000 with an average title length. These lengths are comparable to non-sequels from 2020. Non-sequels from 2010 appear to have the more extreme title lengths. 
